---
title       : googleVis-Tutorial
subtitle    : Based on Gesmann and de Castillo
author      : Vivek Patil
job         : Associate Professor of Marketing
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, interactive]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---


```{r setOptions, message=FALSE,echo=FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')
```

## googleVis

* Provides an interface between Google Chart Tools and R
* Includes the beauty of Hans Rosling's motion charts (purchase of Gapminder by Google)
* https://github.com/mages/googleVis

`install.packages('googleVis')`

Who is Hans Rosling?

---

## Hans Rosling?

9 Ted Talks (through 6/2014... the highest number - next was 6)

<center>
<iframe width="640" height="360" src="http://www.youtube.com/embed/jbkSRLYSojo" frameborder="0" allowfullscreen></iframe>

--- 

## Introduction to Google Chart Tools

* Google Chart Tools provide a way to visualize data on web sites
* The API makes it easy to create interactive charts
* Browser with internet connection required to display chart
* Please read the Google [Terms of Service](https://developers.google.com/terms/) before you start

---

## Key ideas of googleVis

* Transform R data frames into [JSON](http://www.json.org/) objects with [RJSONIO](http://www.omegahat.org/RJSONIO/)

```{r}
library(RJSONIO)
dat <- data.frame(x=LETTERS[1:2], y=1:2)
cat(toJSON(dat)) 
```

* Create wrapper functions in R which generate html files with references to Google's Chart Tools API
* Display the HTML output with the R HTTP help server

---

## The googleVis concept

* Charts: *'gvis' + ChartType*
* For a motion chart we have

```{r eval=FALSE, tidy=FALSE}
M <- gvisMotionChart(data, idvar='id', timevar='date', 
                     options=list(), chartid)
```

* Output of googleVis is a list of list

* Display the chart by simply plotting the output: ```plot(M)```
* Plot will generate a temporary html-file and open it in a new browser window 
* Specific parts can be extracted, e.g. 
  * the chart: ```M$html$chart``` or 
  * data: ```M$html$chart["jsData"]```

---

## Types of charts 

* Histogram, Line, Line with two axes, Bar, Column, Area, Stepped Area, Combo, Scatter, Bubble, Motion, Candlestick, Pie, Gauge, Table, Organization Chart, Tree Map, Sankey, Calendar, and Time line (to check on Annotation, Annotated time line) 
* Mapping: Intensity Map, Geo Chart, Geo markers, Google map markers,Geo Map 

Run ```demo(googleVis)``` to see [examples](http://cran.r-project.org/web/packages/googleVis/vignettes/googleVis_examples.html) of all charts and read the [vignette](http://cran.r-project.org/web/packages/googleVis/vignettes/googleVis.pdf) for more details.

---

## Histogram

```{r results='asis'}
set.seed(123)
datHist=data.frame(A=rpois(100, 20),
                   B=rpois(100, 5),
                   C=rpois(100, 50))

Hist <- gvisHistogram(datHist, options=list(
  legend="{ position: 'top', maxLines: 2 }",
  colors="['#5C3292', '#1A8763', '#871B47']",
  width=750, height=360))
plot(Hist)
```

---

## Line Chart

```{r results='asis'}
df=data.frame(country=c("US", "GB", "BR"), 
              val1=c(10,13,14), 
              val2=c(23,12,32))
Line <- gvisLineChart(df,options=list(width=750, height=500))
plot(Line)
```

---

## Line with two axis

```{r results='asis'}
Line2 <- gvisLineChart(df, "country", c("val1","val2"),
                       options=list(
                         series="[{targetAxisIndex: 0},
                                 {targetAxisIndex:1}]",
                         vAxes="[{title:'val1'}, {title:'val2'}]",
                         width=750, height=400
                       ))
plot(Line2)
```

---

## Customizing Lines

```{r results='asis'}
Dashed <-  gvisLineChart(df, xvar="country", yvar=c("val1","val2"),
                        options=list(
                          series="[{color:'green', targetAxisIndex: 0, 
                          lineWidth: 1, lineDashStyle: [2, 2, 20, 2, 20, 2]}, 
                          {color: 'blue',targetAxisIndex: 1, 
                          lineWidth: 2, lineDashStyle: [4, 1]}]",
                          vAxes="[{title:'val1'}, {title:'val2'}]", width=750,height=400
                        ))
plot(Dashed)
```

---

## Bar Charts

```{r results='asis'}
Bar <- gvisBarChart(df,options=list(width=750, height=400))
plot(Bar)
```

---

## Column Chart

```{r results='asis'}
Column <- gvisColumnChart(df,options=list(width=750, height=400))
plot(Column)
```

---

## Area Chart

```{r results='asis'}
Area <- gvisAreaChart(df,options=list(width=750, height=400))
plot(Area)
```

---

## Stepped Area Chart

```{r results='asis'}
SteppedArea <- gvisSteppedAreaChart(df, xvar="country", 
                                    yvar=c("val1", "val2"),
                                    options=list(isStacked=TRUE, width=750,height=400))
plot(SteppedArea)
```

---

## Combo Chart

```{r results='asis'}
Combo <- gvisComboChart(df, xvar="country",
                        yvar=c("val1", "val2"),
                        options=list(seriesType="bars",
                                     series='{1: {type:"line"}}',width=750,height=400))
plot(Combo)
```

---

## Scatter chart

```{r results='asis'}
Scatter <- gvisScatterChart(women, 
                            options=list(
                              legend="none",
                              pointSize=4,lineWidth=.5, 
                              title="Women", vAxis="{title:'weight (lbs)'}",
                              hAxis="{title:'height (in)'}", 
                              width=700, height=400))
plot(Scatter)
```

---

## Customizing points

```{r results='asis'}
M <- matrix(nrow=6,ncol=6)
M[col(M)==row(M)] <- 1:6
dat <- data.frame(X=1:6, M)
SC <- gvisScatterChart(dat, 
                       options=list(
                         title="Customizing points",
                         legend="right",
                         pointSize=30,
                         series="{
                              0: { pointShape: 'circle' },
                              1: { pointShape: 'triangle' },
                              2: { pointShape: 'square' },
                              3: { pointShape: 'diamond' },
                              4: { pointShape: 'star' },
                              5: { pointShape: 'polygon' }
                              }",
                         width=750,height=400))
```

---
## Customizing points

```{r results='asis'}
plot(SC)
```

---

## More customizations

```{r results='asis'}
Line3 <-  gvisLineChart(df, xvar="country", yvar=c("val1","val2"),
                        options=list(
                          title="Hello World",
                          titleTextStyle="{color:'red', 
                                           fontName:'Courier', 
                                           fontSize:16}",                         
                          backgroundColor="#D3D3D3",                          
                          vAxis="{gridlines:{color:'red', count:3}}",
                          hAxis="{title:'Country', titleTextStyle:{color:'blue'}}",
                          series="[{color:'green', targetAxisIndex: 0}, 
                                   {color: 'orange',targetAxisIndex:1}]",
                          vAxes="[{title:'val1'}, {title:'val2'}]",
                          legend="bottom",
                          curveType="function",
                          width=750,
                          height=400                         
                        ))
```

---
## More customizations

```{r results='asis'}
plot(Line3)
```

---

## Bubble Chart

```{r results='asis'}
Bubble <- gvisBubbleChart(Fruits, idvar="Fruit", 
                          xvar="Sales", yvar="Expenses",
                          colorvar="Year", sizevar="Profit",
                          options=list(
                            hAxis='{minValue:75, maxValue:125}',width=700,height=400))
plot(Bubble)
```

---

## Motion chart example

```{r results='asis'}
M1=gvisMotionChart(Fruits, "Fruit", "Year", 
                     options=list(width=500, height=350))
print(M1,file="M1.html")
```
<iframe src="M1.html" width="600" height="600"></iframe>

---
## Candlestick chart

```{r results='asis'}
Candle <- gvisCandlestickChart(OpenClose, 
                               options=list(legend='none', width=700,height=400))
plot(Candle)
```

---
## Pie charts (for those who insist on not changing) - Not you

```{r results='asis'}
Pie <- gvisPieChart(CityPopularity, options=list(width=700,height=400))
plot(Pie)
```

---
## Gauge charts

```{r results='asis'}
Gauge <-  gvisGauge(CityPopularity, 
                    options=list(min=0, max=800, greenFrom=500,
                                 greenTo=800, yellowFrom=300, yellowTo=500,
                                 redFrom=0, redTo=300, width=400, height=300))
print(Gauge,file="guage.html")
```
<iframe src="guage.html" width="600" height="600"></iframe>

---

## Table

```{r results='asis'}
PopTable <- gvisTable(Population, 
                      formats=list(Population="#,###",
                                   '% of World Population'='#.#%'),
                      options=list(page='enable'))
plot(PopTable)
```

---
## Organization Chart (double-click)

```{r results='asis'}
Org <- gvisOrgChart(Regions, 
                    options=list(width=600, height=250,
                                 size='large', allowCollapse=TRUE))
plot(Org)
```

---
## Tree Map (left/right clicks)

```{r results='asis'}
Tree <- gvisTreeMap(Regions,  
                    "Region", "Parent", 
                    "Val", "Fac", 
                    options=list(fontSize=16))
plot(Tree)
```

---
## Sankey chart

```{r results='asis'}
datSK <- data.frame(From=c(rep("A",3), rep("B", 3)),
                    To=c(rep(c("X", "Y", "Z"),2)),
                    Weight=c(5,7,6,2,9,4))

Sankey <- gvisSankey(datSK, from="From", to="To", weight="Weight",
                     options=list(
                       sankey="{link: {color: { fill: '#d799ae' } },
                            node: { color: { fill: '#a61d4c' },
                            label: { color: '#871b47' } }}"))
print(Sankey,file="Sankey.html")
```

---

## Sankey

<iframe src="Sankey.html" width="600" height="600"></iframe>

---
## Calendar chart

```{r results='asis'}
Cal <- gvisCalendar(Cairo, 
                    datevar="Date", 
                    numvar="Temp",
                    options=list(
                      title="Daily temperature in Cairo",
                      height=320,
                      calendar="{yearLabel: { fontName: 'Times-Roman',
                               fontSize: 32, color: '#1A8763', bold: true},
                               cellSize: 10,
                               cellColor: { stroke: 'red', strokeOpacity: 0.2 },
                               focusedCellColor: {stroke:'red'}}")
)
print(Cal,file="Cal.html")
```

---
## Calendar chart

<iframe src="Cal.html" width="600" height="600"></iframe>


---
## Timeline chart

```{r results='asis'}
datTL <- data.frame(Position=c(rep("President", 3), rep("Vice", 3)),
                    Name=c("Washington", "Adams", "Jefferson",
                           "Adams", "Jefferson", "Burr"),
                    start=as.Date(x=rep(c("1789-03-29", "1797-02-03", 
                                          "1801-02-03"),2)),
                    end=as.Date(x=rep(c("1797-02-03", "1801-02-03", 
                                        "1809-02-03"),2)))

Timeline <- gvisTimeline(data=datTL, 
                         rowlabel="Name",
                         barlabel="Position",
                         start="start", 
                         end="end",
                         options=list(timeline="{groupByRowLabel:false}",
                                      backgroundColor='#ffd', 
                                      height=350,
                                      colors="['#cbb69d', '#603913', '#c69c6e']"))
print(Timeline,file="Timeline.html")
```

---
## Timeline chart

<iframe src="Timeline.html" width="600" height="600"></iframe>


---
## Intensity Map

```{r results='asis'}
Intensity <- gvisIntensityMap(df,options=list( width=750, height=400))
print(Intensity,file="Intensity.html")
```
<iframe src="Intensity.html" width="600" height="600"></iframe>

---
## Geo Chart

```{r results='asis'}
Geo=gvisGeoChart(Exports, locationvar="Country", 
                 colorvar="Profit",
                 options=list(projection="kavrayskiy-vii",
                 width=750, height=400))
print(Geo,file="Geo.html")
```
<iframe src="Geo.html" width="600" height="600"></iframe>

---
## Geo chart US

```{r results='asis'}
library(datasets)
states <- data.frame(state.name, state.x77)
GeoStates <- gvisGeoChart(states, "state.name", "Illiteracy",
                          options=list(region="US", displayMode="regions", resolution="provinces",
                                       width=600, height=400))
print(GeoStates,file="GeoStates.html")
```
<iframe src="GeoStates.html" width="600" height="600"></iframe>

---
## Geo markers: Hurricane Andrew track

```{r results='asis'}
GeoMarker <- gvisGeoChart(Andrew, "LatLong", 
                          sizevar='Speed_kt',
                          colorvar="Pressure_mb", 
                          options=list(region="US"))
print(GeoMarker,file="GeoMarker.html")
```
<iframe src="GeoMarker.html" width="600" height="600"></iframe>

---
## Google map: Hurricane Andrew Track

```{r results='asis'}
AndrewMap <- gvisMap(Andrew, "LatLong" , "Tip", 
                     options=list(showTip=TRUE, 
                                  showLine=TRUE, 
                                  enableScrollWheel=TRUE,
                                  mapType='terrain', 
                                  useMapTypeControl=TRUE))
print(AndrewMap,file="AndrewMap.html")
```
<iframe src="AndrewMap.html" width="600" height="600"></iframe>

---
## Geo Map: Hurricane Andrew Track

```{r results='asis'}
AndrewGeo <- gvisGeoMap(Andrew, 
                        locationvar="LatLong", 
                        numvar="Speed_kt", 
                        hovervar="Category", 
                        options=list(height=350, 
                                     region="US", 
                                     dataMode="markers"))
print(AndrewGeo,file="AndrewGeo.html")
```
<iframe src="AndrewGeo.html" width="600" height="600"></iframe>

---
## Merging

```{r results='asis'}
G <- gvisGeoChart(Exports, "Country", "Profit", 
                  options=list(width=600, height=400))
T <- gvisTable(Exports, 
               options=list(width=440, height=400))

GT <- gvisMerge(G,T, horizontal=TRUE) 
print(GT,file="GT.html")
```
<iframe src="GT.html" width="600" height="600"></iframe>

---

## Your turn to have fun!!

Code for this deck can be found at: https://github.com/patilv/googleVis-tutorial

